Discriminative Sensing Based on Signal Processing for Information Security Analysis

### 2017-10-27 docs/thesis/chapters/3_mobile/mobile.tex
introducion...
conclusion and introduction

\section{Introduction}
\label{sec_intro}
Cloud computing is a new rapidly evolving paradigm in the world of distributed networking and computation. The basic features of the cloud environment is providing the elastic, on-demand and secure services for the end-users. While the first two requirements are rather well conceptualized and supported by the majority of the cloud platforms in use, security is a serious concern of the cloud providers and governmental organizations as well as academia and research community \cite{csa2016,higashi2015,gartner2015}. Although for the small and medium-sized enterprises (SME) the cloud environment is often the most cost-effective and easily scalable solution, the security and privacy of the sensitive data in cloud platforms are also not fully conceptualized leading to obscure and incomplete security paradigms and solutions.

@@

The paper is structured as follows. Section \ref{sec_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions for offline protection in the BYOD world. Section \ref{sec_the_client_security_architecture} outlines the mobile security architecture of the proposed solution. Section \ref{sec_offline_mode} presents the detailed scheme of the proposed solution to the problem of offline mobile client security. Section \ref{sec_algorithms} explains the algorithms for the core security methods: AES, ABE, SSS and MOS. Section \ref{sec_results} discusses the security proofs including the outline of adversary model, the common threat scenarios, the data modeling, the complexity analysis and dicussion of the practical implementation results. Section \ref{sec_conclusion} concludes the paper.

\section{Related works}
\label{sec_related_works}
The increasing usage of BYOD demands more sophisticated data protection services compared to ordinary computing environment. A common practice is to provide additional contextual methods apart from authentication, DLP services, and encryption, which can be at rest, in transit and in use \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. The contextual methods increase the security of the mobile client at a maximum level with minimum resource requirements. The most commonly used are:

@@

\section{The mobile security architecture}
\label{sec_the_client_security_architecture}
The approach proposed in this work describes and implements the complete lifecycle of the mobile client security infrastructure. The mobile security processes depend on the key expiry period, and are used to access the protected storage. Once the user keys expire, the user is requested to enter his valid credentials, i.e. PIN and password. The mobile client then sends the credentials to the server for verification. Once the new set of access keys is received, the user can access the protected files in the offline mode, without the access to the server. This means that no further communication with the server is needed until the key expires.

The core set of functions and protocols can be divided into three sets of operations as shown in Figure \ref{fig:01}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=5cm]{fig01.eps}
	\caption{The core set of functions and protocols of the mobile cloud security infrastructure}
	\label{fig:01}
\end{figure}

Figure \ref{fig:01} describes the mobile client protection both in online and offline mode. In online mode, the client has the possibility to connect to the server and the security of the client is enhanced by the server-backed up mechanisms. On the other hand, in offline mode the client’s security is supported by the standalone mechanisms. Additionally, the mobile client protection is enhanced by the threat intelligence unit providing the constant monitoring and analysis.

Figure \ref{fig:02} depicts the client-side protection mechanisms. The client should support 4 subsystems: i) encryption subsystem that contains the procedures of encryption and decryption; ii) storage subsystem that provides the downloaded shares and key storage protection; iii) threat intelligence unit that provides the constant monitoring, and iv) the communication subsystem. In short, all security procedures are connected to 4 groups of operations: file request and receiving; encryption and decryption; file and key storing; monitoring and analysis.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig02.eps}
	\caption{The Mobile Client Architecture}
	\label{fig:02}
\end{figure}

@@

The cryptographic functions include the decryption (for AES and ABE) key restoration for SSS and EKE token generation procedure. While the AES-ABE hybrid encryption scheme functions both in the online and offline modes, the SSS is used only in the offline mode, while the EKE generation and executions is used in the online mode. These operations are described in Subsections \ref{sec_aes}-\ref{sec_secret} and are part of the Figure \ref{fig:01}.

The threat intelligence infrastructure takes into account simple actors such as the time counter for the key expiry period, the counter of unsuccessful tries in order to protect from brute force attacks, and MOS-inspired statistics analyzer. Functions such as alerting and deleting the expired key belong to this block as well. These functions are described in Subsection \ref{sec_mos}.

@@

\section{The proposed solution for offline mobile security}
\label{sec_offline_mode}
This section proposes an approach for the mobile client protection in which the security is supported both in online and offline modes. Currently and to the best of our knowledge, the systems of mobile client protection follow a model where the protected client can operate only when it is connected to the cloud, which is not always convenient for the end-user. The basic principles of the mobile client protection herein proposed are:

@@

The proposed architecture for offline mobile security is represented in Figure \ref{fig:03}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig03.eps}
	\caption{Proposed Architecture for Offline Mobile Security}
	\label{fig:03}
\end{figure}

The Figure \ref{fig:03} describes the proposed architecture for offline mobile security, showing the modules responsible for securing the mobile client, which includes:

\begin{enumerate}
	\item \textbf{Protected Storage}: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files. The outline of the hybrid encryption scheme is presented in the Section \ref{sec_algorithms}.
	\item \textbf{Threat Intelligence Manager (TIM)}: most attacks incur into significant variation on the legitimate behavior of information systems, or they adopt well-known patterns that can be easily detected by monitoring the system in the case of the offline mode. Signal processing techniques have been successfully applied to anomaly detection \cite{lu2009network, huang2009signal} and have become a solution to a problem of improving detection accuracy, adaptability and computational cost for application on resource-constrained scenarios. Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior, according to the scenarios in Section \ref{sec_results}. Moreover, Model Order Selection (MOS), which is an effective signal processing technique to separate noise components from the principal components, can be applied into anomaly and attack detection \cite{tenorio2013greatest}, to identify and separate malicious behaviors from the legitimate ones. The TIM is an internal module of the mobile client that implements offline anomaly detection through signal processing techniques.
	\item \textbf{Key Management Center}: it includes the functions for maintaining the key expiry period and deleting the expired keys. 
\end{enumerate}

\subsection{Offline mode workflow}
\label{sec_offline_mode_workflow}

@@

The mobile client performs all the cryptographic calculations in the shadow. These calculations include the key storage, key restoring, decryption and showing the decrypted files in the mobile client area. Note that the mobile client does not check the password or PIN validity as it does not hold the verification proof for the above. Also, the mobile client keeps track of the user activities and the key expiry. The complete list of the mobile client activities is presented in the Figure \ref{fig:04}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig04.eps}
	\caption{Offline Mode Operations}
	\label{fig:04}
\end{figure}

To finalize the description of the offline mode, Figure \ref{fig:05} shows the complete workflow of the proposed mobile application in the offline-mode.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig05.eps}
	\caption{Offline mode authorization workflow}
	\label{fig:05}
\end{figure}

The core cryptographic module of the implemented solution is based on the combination of AES-ABE-SSS methods, described in the Section \ref{sec_algorithms}. The key feature of the offline security is that the mobile client does not actually store any part of the user password to be verified. The mobile client combines its own key with the user share (PIN and password-derived) in order to restore the initial KEY\_SET\_KEY. If the user provides the wrong share the mobile client will not be able to recognize it, but will decrypt the files incorrectly. Additionally, the password entering is tracked and too many tries in a short time are considered a threat.

\subsection{Offline Behavioral Analysis}
\label{sec_offline_behavioral_analysis}

In the proposed client security architecture, the Threat Intelligence Manager (TIM) is responsible for receiving logged user operations, perform feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode. Figure \ref{fig:06} depicts the TIM for offline behavioral analysis.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig06.eps}
	\caption{The Threat Intelligence Manager Workflow}
	\label{fig:06}
\end{figure}

As depicted in Figure \ref{fig:06}, users request operations are logged so that the main features can be monitored in the mobile client. The user behavior, trying or effectively executing operations, shall be incrementally captured and logged, making possible to monitor the main features that can reveal malicious behaviors, as well as to identify unexpected behaviors that can reveal possible threats. Therefore, the user operations are monitored by the client app, which sends the information to the User Activity Logging (UAL).

@@

The LAC performs the behavioral analysis through eigenvalue analysis and MOS schemes, which identify anomalies on sparse, subtle or abrupt number of user operations. The malicious behavior detection is detail described in \ref{sec_mos}.

\section{The algorithms, key usage and data protection methods}
\label{sec_algorithms}
In the proposed approach, the kernel encryption scheme in the mobile client is a combination of several methods of security. The files are encrypted with 128/256-bit AES, while the permanent file keys are encrypted with the attribute-based encryption. The set of expiring ABE keys corresponding to the set of files accessible by user in encrypted with a single expiring AES key (KEY\_SET\_KEY). This key is expiring and is splited by the server into 4 parts (2 are stored on the device and 2 belong to the user) by the method of secure secret sharing. The encryption workflow is outlined in the Figure \ref{fig:07}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig07.eps}
	\caption{Encryption workflow}
	\label{fig:07}
\end{figure}

The key hierarchy is explained by Figure \ref{fig:08} where the key represented in Figure \ref{fig:07} serves for decrypting the key for each level represented in Figure \ref{fig:08}. As it is possible to be observed in Figure \ref{fig:08}, the master key on the upper level is computed by means of SSS (Subsection \ref{sec_secret}). It serves for decrypting the ABE key storage (section \ref{sec_abe}). The ABE key allows the user finally decrypt the permanent file share keys. Further, the file share can be decrypted by means of AES.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig08.eps}
	\caption{Encryption key hierarchy}
	\label{fig:08}
\end{figure}

@@

In Subsections \ref{sec_aes}-\ref{sec_secret}, a detailed description of the building blocks of the hybrid encryption is provided, and the proposed malicious behavior detection scheme is detail described in \ref{sec_mos}.

\subsection{AES file encryption }
\label{sec_aes}
The protected files are encrypted on the server side with the secure 128/256 bit AES encryption, which is currently an industrial standard. Nevertheless, it is possible to encrypt the file storage with Blowfish or Serpent which also provide a high level of safety \cite{nist2000}.

@@

The scheme of the server-side AES encryption is represented in Figure \ref{fig:09}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{fig09.eps}
	\caption{Server-side encryption}
	\label{fig:09}
\end{figure}

@@

\subsection{ABE encryption to protect the FILE\_KEY }
\label{sec_abe}
With each user session, the permanent FILE\_KEY (uni\-que AES key) is re-encrypted. The set of FILE\_KEYs is protected with the corresponding ABE keys. The unique ABE model that we propose supports the attribute policy based on user groups and on file shares, i.e. the key attributes correspond both to the groups and the file shares.

@@

\begin{equation}\label{eq:eq00}
Y^s = e(g,g)^{ys} = e(E_i,D_i) = e(g^{\frac{yw}{t_i}},g^{\frac{t_is}{w}})=e(g,g)^{ys}.
\end{equation}

@@

\subsection{Secret sharing scheme to protect the key storage }
\label{sec_secret}

@@

\begin{equation}\label{eq:eq01}
  \left\{
  \begin{array}{@{}ll@{}}
    S = s \textrm{ mod } p_0
    \\
    S = s_1 \textrm{ mod } p_1
  \end{array}\right.
\end{equation}

According to the CRT:
\begin{equation}\label{eq:eq02}
S \equiv {s_{1}p_{0}{p_0}^{-1}}_{p1} + {s_{1}p_{1}{p_1}^{-1}}_{p0} + \textrm{ mod } p_0p_1
\end{equation}

@@

\subsection{MOS for Threat Intelligence}
\label{sec_mos}
In the context of anomaly-based schemes for attack detection, the proposed behavioral analysis approach applies signal processing techniques, such as Principal Component Analysis and Model Order Selection schemes \cite{tenorio2013greatest}, for automatic detection of attacks or malicious behaviors.

@@

\begin{equation}\label{eq:eq03}
\mathbf{y}_{m}^{(q)} = \mathbf{x}_{m}^{(q)} - \bar{\mathbf{x}}_{m}^{(q)}.
\end{equation}

@@

\begin{equation}\label{eq:eq04}
\mathbf{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\mathbf{Y}^{(q)}\mathbf{Y}^{(q)^{\rm T}},
\end{equation}
where $\mathbf{\hat{R}}_{yy}^{(q)}$ means the estimation of the sample covariance matrix from the measured zero mean matrix $\mathbf{Y}^{(q)}$ over $N$ minutes of the time frame $q$.

@@

\begin{equation}\label{eq:eq05}
\mathbf{z}_{m}^{(q)} = \frac{\mathbf{x}_{m}^{(q)} - \bar{\mathbf{x}}_{m}^{(q)}}{\mathbf{\sigma}_{m}^{(q)}}.
\end{equation}

@@

\begin{equation}\label{eq:eq06}
\mathbf{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\mathbf{Z}^{(q)}\mathbf{Z}^{(q)^{\rm T}}.
\end{equation}

@@

\begin{equation}\label{eq:eq07}
\mathbf{e}^{(q)} = \rm diag(\mathbf{\Lambda}^{(q)}),
\end{equation}

@@

\begin{equation}\label{eq:eq08}
\mathbf{E} =
\begin{bmatrix}
  \lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
  \lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
  \lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
  \vdots & \vdots & \ddots & \vdots  \\
  \lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
\end{bmatrix}.
\end{equation}

@@

\begin{equation}\label{eq:eq09}
\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\} = [ \lambda_1^{(1)}, \lambda_1^{(2)} \cdots \lambda_1^{(Q)}]
\end{equation}

@@

\begin{equation}\label{eq:eq09}
\hat{d} = \textrm{MOS}(\boldsymbol{e}_{\rm max})
\end{equation}

@@

\section{Results and analysis}
\label{sec_results}

@@

\subsection{Security analysis}
\label{sec_sec_analysis}
The security analysis of the proposed model was performed both from the point of implemented cryptographic mechanisms and from the user behavioral analysis. Two common attack scenarios were analyzed. First, the malicious outsider trying to infect or steal the important files. Second, the malicious expired user trying to steal the important files.

\subsubsection{Adversary model}
\label{sec_adversary_model}
The proposal is to use a restricted adversary model. In other words, the actions of the adversary are predicted within the outlines model and scenarios. For example, it is not possible to guarantee the security of the decrypted document once the user copies its content to another file.

@@

\subsubsection{Security analysis}
\label{sec_sec_analysis}

In the offline mode the mobile client does not store the user password (see Section \ref{sec_offline_mode_workflow} for details), i.e. no information about the password leaks, and therefore there is no possibility for the malicious user to check if the password he is trying to enter is correct or not. The possible scenario for the information leakage in the case of the malicious outsider is if:

@@

\subsubsection{Common threat scenarios}
\label{sec_common}
This section provides the detailed description of the common scenarios in which the log and behavioral analysis is provided. The behavioral analysis can help to keep the user or administrator informed of the threat and take actions, as well as it can be useful in order to implement threat preventions or reactive actions to avoid threat propagation.

@@

\subsubsection{Data Modeling for Behavioral Analysis }
\label{sec_data}

MOS schemes are used in order to identify anomalous behavior that can indicate an attack and be used to prevent or avoid attack propagation. Therefore, it is necessary to analyze the data that can be collected from user operations on mobile client, to identify features that can be modeled and submitted to MOS schemes, according to described in Section \ref{sec_mos}.

@@

\begin{table*}[!t]
	\caption{Speed of decryption on a client device}
	\label{tab:01}
	\centering
	\begin{tabular}{|l|l|l|l|l|l|}
		\hline \rowcolor{Gray} Device Model	& File set size	& \begin{tabular}[x]{@{}l@{}}Mean time of\\getting decrypted\\stream (ABE)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Mean time of\\decrypting \\content (AES)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Mean time of\\copying file\\without decrypt\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Mean time of\\decryption\\(the whole file)\end{tabular}	\\ \hline

@@

Table \ref{tab:01} presents the results of mobile client security workflow testing and demonstrates that the proposed scheme decrypts the data with acceptable speed. One can compare the speed of ABE/AES decryption (keys and actual data) and the speed of the whole decryption process.

Table \ref{tab:02} demonstrates the results of testing the speed of the mobile client for receiving and displaying the list of files. The client was tested for displaying up to 1000 files. Around 100 requests were generated from 10 threads. It is possible to observe that the client receives and displays the list of files depending on the user attributes with acceptable average speed, showing that the proposed protection scheme is usable.

>>>>

\section{Introduction}
\label{sec:3_intro}
Cloud computing is a new rapidly evolving paradigm in the world of distributed networking and computation. The basic features of the cloud environment is providing the elastic, on-demand and secure services for the end-users. While the first two requirements are rather well conceptualized and supported by the majority of the cloud platforms in use, security is a serious concern of the cloud providers and governmental organizations as well as academia and research community \cite{csa2016,higashi2015,gartner2015}. Although for the small and medium-sized enterprises (SME) the cloud environment is often the most cost-effective and easily scalable solution, the security and privacy of the sensitive data in cloud platforms are also not fully conceptualized leading to obscure and incomplete security paradigms and solutions.

@@

The paper is structured as follows. Section \ref{sec:3_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions for offline protection in the BYOD world. Section \ref{sec:3_the_client_security_architecture} outlines the mobile security architecture of the proposed solution. Section \ref{sec:3_offline_mode} presents the detailed scheme of the proposed solution to the problem of offline mobile client security. Section \ref{sec:3_algorithms} explains the algorithms for the core security methods: AES, ABE, SSS and MOS. Section \ref{sec:3_results} discusses the security proofs including the outline of adversary model, the common threat scenarios, the data modeling, the complexity analysis and dicussion of the practical implementation results. Section \ref{sec:3_conclusion} concludes the paper.

\section{Related works}
\label{sec:3_related_works}
The increasing usage of BYOD demands more sophisticated data protection services compared to ordinary computing environment. A common practice is to provide additional contextual methods apart from authentication, DLP services, and encryption, which can be at rest, in transit and in use \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. The contextual methods increase the security of the mobile client at a maximum level with minimum resource requirements. The most commonly used are:

@@

\section{The mobile security architecture}
\label{sec:3_the_client_security_architecture}
The approach proposed in this work describes and implements the complete lifecycle of the mobile client security infrastructure. The mobile security processes depend on the key expiry period, and are used to access the protected storage. Once the user keys expire, the user is requested to enter his valid credentials, i.e. PIN and password. The mobile client then sends the credentials to the server for verification. Once the new set of access keys is received, the user can access the protected files in the offline mode, without the access to the server. This means that no further communication with the server is needed until the key expires.

The core set of functions and protocols can be divided into three sets of operations as shown in Figure \ref{fig:3_01}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=8cm]{figures/fig01.eps}
	\caption{The core set of functions and protocols of the mobile cloud security infrastructure}
	\label{fig:3_01}
\end{figure}

Figure \ref{fig:3_01} describes the mobile client protection both in online and offline mode. In online mode, the client has the possibility to connect to the server and the security of the client is enhanced by the server-backed up mechanisms. On the other hand, in offline mode the client’s security is supported by the standalone mechanisms. Additionally, the mobile client protection is enhanced by the threat intelligence unit providing the constant monitoring and analysis.

Figure \ref{fig:3_02} depicts the client-side protection mechanisms. The client should support 4 subsystems: i) encryption subsystem that contains the procedures of encryption and decryption; ii) storage subsystem that provides the downloaded shares and key storage protection; iii) threat intelligence unit that provides the constant monitoring, and iv) the communication subsystem. In short, all security procedures are connected to 4 groups of operations: file request and receiving; encryption and decryption; file and key storing; monitoring and analysis.

\begin{figure}[h!]
	\centering
	\includegraphics[width=12cm]{figures/fig02.eps}
	\caption{The Mobile Client Architecture}
	\label{fig:3_02}
\end{figure}

@@

The cryptographic functions include the decryption (for AES and ABE) key restoration for SSS and EKE token generation procedure. While the AES-ABE hybrid encryption scheme functions both in the online and offline modes, the SSS is used only in the offline mode, while the EKE generation and executions is used in the online mode. These operations are described in Subsections \ref{sec:3_aes}-\ref{sec:3_secret} and are part of the Figure \ref{fig:3_01}.

The threat intelligence infrastructure takes into account simple actors such as the time counter for the key expiry period, the counter of unsuccessful tries in order to protect from brute force attacks, and MOS-inspired statistics analyzer. Functions such as alerting and deleting the expired key belong to this block as well. These functions are described in Subsection \ref{sec:3_mos}.

@@

\section{The proposed solution for offline mobile security}
\label{sec:3_offline_mode}
This section proposes an approach for the mobile client protection in which the security is supported both in online and offline modes. Currently and to the best of our knowledge, the systems of mobile client protection follow a model where the protected client can operate only when it is connected to the cloud, which is not always convenient for the end-user. The basic principles of the mobile client protection herein proposed are:

@@

The proposed architecture for offline mobile security is represented in Figure \ref{fig:3_03}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=15cm]{figures/fig03.eps}
	\caption{Proposed Architecture for Offline Mobile Security}
	\label{fig:3_03}
\end{figure}

The Figure \ref{fig:3_03} describes the proposed architecture for offline mobile security, showing the modules responsible for securing the mobile client, which includes:

\begin{enumerate}
	\item \textbf{Protected Storage}: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files. The outline of the hybrid encryption scheme is presented in the Section \ref{sec:3_algorithms}.
	\item \textbf{Threat Intelligence Manager (TIM)}: most attacks incur into significant variation on the legitimate behavior of information systems, or they adopt well-known patterns that can be easily detected by monitoring the system in the case of the offline mode. Signal processing techniques have been successfully applied to anomaly detection \cite{lu2009network, huang2009signal} and have become a solution to a problem of improving detection accuracy, adaptability and computational cost for application on resource-constrained scenarios. Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior, according to the scenarios in Section \ref{sec:3_results}. Moreover, Model Order Selection (MOS), which is an effective signal processing technique to separate noise components from the principal components, can be applied into anomaly and attack detection \cite{tenorio2013greatest}, to identify and separate malicious behaviors from the legitimate ones. The TIM is an internal module of the mobile client that implements offline anomaly detection through signal processing techniques.
	\item \textbf{Key Management Center}: it includes the functions for maintaining the key expiry period and deleting the expired keys. 
\end{enumerate}

\subsection{Offline mode workflow}
\label{sec:3_offline_mode_workflow}

@@

The mobile client performs all the cryptographic calculations in the shadow. These calculations include the key storage, key restoring, decryption and showing the decrypted files in the mobile client area. Note that the mobile client does not check the password or PIN validity as it does not hold the verification proof for the above. Also, the mobile client keeps track of the user activities and the key expiry. The complete list of the mobile client activities is presented in the Figure \ref{fig:3_04}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=12cm]{figures/fig04.eps}
	\caption{Offline Mode Operations}
	\label{fig:3_04}
\end{figure}

To finalize the description of the offline mode, Figure \ref{fig:3_05} shows the complete workflow of the proposed mobile application in the offline-mode.

\begin{figure}[h!]
	\centering
	\includegraphics[width=12cm]{figures/fig05.eps}
	\caption{Offline mode authorization workflow}
	\label{fig:3_05}
\end{figure}

The core cryptographic module of the implemented solution is based on the combination of AES-ABE-SSS methods, described in the Section \ref{sec:3_algorithms}. The key feature of the offline security is that the mobile client does not actually store any part of the user password to be verified. The mobile client combines its own key with the user share (PIN and password-derived) in order to restore the initial KEY\_SET\_KEY. If the user provides the wrong share the mobile client will not be able to recognize it, but will decrypt the files incorrectly. Additionally, the password entering is tracked and too many tries in a short time are considered a threat.

\subsection{Offline Behavioral Analysis}
\label{sec:3_offline_behavioral_analysis}

In the proposed client security architecture, the Threat Intelligence Manager (TIM) is responsible for receiving logged user operations, perform feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode. Figure \ref{fig:3_06} depicts the TIM for offline behavioral analysis.

\begin{figure}[h!]
	\centering
	\includegraphics[width=10cm]{figures/fig06.eps}
	\caption{The Threat Intelligence Manager Workflow}
	\label{fig:3_06}
\end{figure}

As depicted in Figure \ref{fig:3_06}, users request operations are logged so that the main features can be monitored in the mobile client. The user behavior, trying or effectively executing operations, shall be incrementally captured and logged, making possible to monitor the main features that can reveal malicious behaviors, as well as to identify unexpected behaviors that can reveal possible threats. Therefore, the user operations are monitored by the client app, which sends the information to the User Activity Logging (UAL).

@@

The LAC performs the behavioral analysis through eigenvalue analysis and MOS schemes, which identify anomalies on sparse, subtle or abrupt number of user operations. The malicious behavior detection is detail described in \ref{sec:3_mos}.

\section{The algorithms, key usage and data protection methods}
\label{sec:3_algorithms}
In the proposed approach, the kernel encryption scheme in the mobile client is a combination of several methods of security. The files are encrypted with 128/256-bit AES, while the permanent file keys are encrypted with the attribute-based encryption. The set of expiring ABE keys corresponding to the set of files accessible by user in encrypted with a single expiring AES key (KEY\_SET\_KEY). This key is expiring and is splited by the server into 4 parts (2 are stored on the device and 2 belong to the user) by the method of secure secret sharing. The encryption workflow is outlined in the Figure \ref{fig:3_07}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=12cm]{figures/fig07.eps}
	\caption{Encryption workflow}
	\label{fig:3_07}
\end{figure}

The key hierarchy is explained by Figure \ref{fig:3_08} where the key represented in Figure \ref{fig:3_07} serves for decrypting the key for each level represented in Figure \ref{fig:3_08}. As it is possible to be observed in Figure \ref{fig:3_08}, the master key on the upper level is computed by means of SSS (Subsection \ref{sec:3_secret}). It serves for decrypting the ABE key storage (section \ref{sec:3_abe}). The ABE key allows the user finally decrypt the permanent file share keys. Further, the file share can be decrypted by means of AES.

\begin{figure}[h!]
	\centering
	\includegraphics[width=10cm]{figures/fig08.eps}
	\caption{Encryption key hierarchy}
	\label{fig:3_08}
\end{figure}

@@

In Subsections \ref{sec:3_aes}-\ref{sec:3_secret}, a detailed description of the building blocks of the hybrid encryption is provided, and the proposed malicious behavior detection scheme is detail described in \ref{sec:3_mos}.

\subsection{AES file encryption }
\label{sec:3_aes}
The protected files are encrypted on the server side with the secure 128/256 bit AES encryption, which is currently an industrial standard. Nevertheless, it is possible to encrypt the file storage with Blowfish or Serpent which also provide a high level of safety \cite{nist2000}.

@@

The scheme of the server-side AES encryption is represented in Figure \ref{fig:3_09}.

\begin{figure}[h!]
	\centering
	\includegraphics[width=12cm]{figures/fig09.eps}
	\caption{Server-side encryption}
	\label{fig:3_09}
\end{figure}

@@

\subsection{ABE encryption to protect the FILE\_KEY }
\label{sec:3_abe}
With each user session, the permanent FILE\_KEY (uni\-que AES key) is re-encrypted. The set of FILE\_KEYs is protected with the corresponding ABE keys. The unique ABE model that we propose supports the attribute policy based on user groups and on file shares, i.e. the key attributes correspond both to the groups and the file shares.

@@

\begin{equation}\label{eq:3_eq00}
Y^s = e(g,g)^{ys} = e(E_i,D_i) = e(g^{\frac{yw}{t_i}},g^{\frac{t_is}{w}})=e(g,g)^{ys}.
\end{equation}

@@

\subsection{Secret sharing scheme to protect the key storage }
\label{sec:3_secret}

@@

\begin{equation}\label{eq:3_eq01}
  \left\{
  \begin{array}{@{}ll@{}}
    S = s \textrm{ mod } p_0
    \\
    S = s_1 \textrm{ mod } p_1
  \end{array}\right.
\end{equation}

According to the CRT:
\begin{equation}\label{eq:3_eq02}
S \equiv {s_{1}p_{0}{p_0}^{-1}}_{p1} + {s_{1}p_{1}{p_1}^{-1}}_{p0} + \textrm{ mod } p_0p_1
\end{equation}

@@

\subsection{MOS for Threat Intelligence}
\label{sec:3_mos}
In the context of anomaly-based schemes for attack detection, the proposed behavioral analysis approach applies signal processing techniques, such as Principal Component Analysis and Model Order Selection schemes \cite{tenorio2013greatest}, for automatic detection of attacks or malicious behaviors.

@@

\begin{equation}\label{eq:3_eq03}
\mathbf{y}_{m}^{(q)} = \mathbf{x}_{m}^{(q)} - \bar{\mathbf{x}}_{m}^{(q)}.
\end{equation}

@@

\begin{equation}\label{eq:3_eq04}
\mathbf{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\mathbf{Y}^{(q)}\mathbf{Y}^{(q)^{\rm T}},
\end{equation}
where $\mathbf{\hat{R}}_{yy}^{(q)}$ means the estimation of the sample covariance matrix from the measured zero mean matrix $\mathbf{Y}^{(q)}$ over $N$ minutes of the time frame $q$.

@@

\begin{equation}\label{eq:3_eq05}
\mathbf{z}_{m}^{(q)} = \frac{\mathbf{x}_{m}^{(q)} - \bar{\mathbf{x}}_{m}^{(q)}}{\mathbf{\sigma}_{m}^{(q)}}.
\end{equation}

@@

\begin{equation}\label{eq:3_eq06}
\mathbf{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\mathbf{Z}^{(q)}\mathbf{Z}^{(q)^{\rm T}}.
\end{equation}

@@

\begin{equation}\label{eq:3_eq07}
\mathbf{e}^{(q)} = \rm diag(\mathbf{\Lambda}^{(q)}),
\end{equation}

@@

\begin{equation}\label{eq:3_eq08}
\mathbf{E} =
\begin{bmatrix}
  \lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
  \lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
  \lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
  \vdots & \vdots & \ddots & \vdots  \\
  \lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
\end{bmatrix}.
\end{equation}

@@

\begin{equation}\label{eq:3_eq09}
\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\} = [ \lambda_1^{(1)}, \lambda_1^{(2)} \cdots \lambda_1^{(Q)}]
\end{equation}

@@

\begin{equation}\label{eq:3_eq10}
\hat{d} = \textrm{MOS}(\boldsymbol{e}_{\rm max})
\end{equation}

@@

\section{Results and analysis}
\label{sec:3_results}

@@

\subsection{Security analysis}
\label{sec:3_sec_analysis}
The security analysis of the proposed model was performed both from the point of implemented cryptographic mechanisms and from the user behavioral analysis. Two common attack scenarios were analyzed. First, the malicious outsider trying to infect or steal the important files. Second, the malicious expired user trying to steal the important files.

\subsubsection{Adversary model}
\label{sec:3_adversary_model}
The proposal is to use a restricted adversary model. In other words, the actions of the adversary are predicted within the outlines model and scenarios. For example, it is not possible to guarantee the security of the decrypted document once the user copies its content to another file.

@@

\subsubsection{Security analysis}
\label{sec:3_sec_analysis2}

In the offline mode the mobile client does not store the user password (see Section \ref{sec:3_offline_mode_workflow} for details), i.e. no information about the password leaks, and therefore there is no possibility for the malicious user to check if the password he is trying to enter is correct or not. The possible scenario for the information leakage in the case of the malicious outsider is if:

@@

\subsubsection{Common threat scenarios}
\label{sec:3_common}
This section provides the detailed description of the common scenarios in which the log and behavioral analysis is provided. The behavioral analysis can help to keep the user or administrator informed of the threat and take actions, as well as it can be useful in order to implement threat preventions or reactive actions to avoid threat propagation.

@@

\subsubsection{Data Modeling for Behavioral Analysis }
\label{sec:3_data}

MOS schemes are used in order to identify anomalous behavior that can indicate an attack and be used to prevent or avoid attack propagation. Therefore, it is necessary to analyze the data that can be collected from user operations on mobile client, to identify features that can be modeled and submitted to MOS schemes, according to described in Section \ref{sec:3_mos}.

@@

\begin{table*}[!t]
	\caption{Speed of decryption on a client device}
	\scriptsize
	\label{tab:3_01}
	\centering
	\begin{tabular}{|l|l|l|l|l|l|}
		\hline \rowcolor{Gray} Device Model	& File set size	& \begin{tabular}[x]{@{}l@{}}Mean time of\\getting decrypted\\stream (ABE)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Mean time of\\decrypting \\content (AES)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Mean time of\\copying file\\without decrypt\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Mean time of\\decryption\\(the whole file)\end{tabular}	\\ \hline

@@

Table \ref{tab:3_01} presents the results of mobile client security workflow testing and demonstrates that the proposed scheme decrypts the data with acceptable speed. One can compare the speed of ABE/AES decryption (keys and actual data) and the speed of the whole decryption process.

Table \ref{tab:3_02} demonstrates the results of testing the speed of the mobile client for receiving and displaying the list of files. The client was tested for displaying up to 1000 files. Around 100 requests were generated from 10 threads. It is possible to observe that the client receives and displays the list of files depending on the user attributes with acceptable average speed, showing that the proposed protection scheme is usable.
